---
title: "Tidytuesday 16/11/2021"
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(patchwork)
```

## Get the data

population density per km2

```{r}
# raster objects
# turn raster objects to data frames and return the x and y coordinates with xy = tRUE
# rename 3rd column to be population
pop2000 <- afrilearndata::afripop2000 %>% raster::as.data.frame(xy = TRUE) %>% rename(population = 3) %>% filter(!is.na(population))
pop2020 <- afrilearndata::afripop2020 %>% raster::as.data.frame(xy = TRUE) %>% rename(population = 3) %>% filter(!is.na(population))
```

## The question

How did population density change between 2000 and 2020? Which places were popular to move to?

```{r}
ggplot(pop2000)+
  geom_tile(aes(x=x, y=y, fill = population))+
  theme_void()
```

Looks like there are a lot of low numbers.
Maybe log transform??

```{r}
ggplot(pop2000)+
  geom_tile(aes(x=x, y=y, fill = log(population)))+
  scale_fill_viridis_c() +
  theme_void()
```

# other approach

convert high values to see differences better

Also not very helpful... Look at the histogram of population densities

```{r}
ggplot(pop2000) + 
  geom_histogram(aes(x=population))

ggplot(pop2020) +
  geom_histogram(aes(x=population))
```

Most population densities are very low...

Which threshold can be chosen for plotting?

Idea from https://github.com/gkaramanis/tidytuesday/blob/master/2021/2021-week45/afrimapr.R

```{r}
pop2000 %>% 
  filter(population < 300) %>% 
ggplot() + 
  geom_histogram(aes(x=population))
```


Relable everything higher than the threshold

```{r}
threshold = 500

pop2000_final <- pop2000 %>% 
  mutate(population = case_when(
    population <= threshold ~ population,
    TRUE ~ threshold + 100
  ))
pop2020_final <- pop2020 %>% 
  mutate(population = case_when(
    population <= threshold ~ population,
    TRUE ~ threshold + 100
  ))
```


Make the same plot again

```{r}
# library(showtext)
# font_paths()
# font_files() # available fonts
# font_families()
# 
# showtext_auto()
# 
# sysfonts::font_add_google(name = "Architects Daughter")

pop2000_plot <- 
ggplot(pop2000_final)+
  geom_tile(aes(x=x, y=y, fill = population))+
  ggtitle("2020")

pop2020_plot <- 
ggplot(pop2020_final)+
  geom_tile(aes(x=x, y=y, fill = population))+
  ggtitle("2020")

# color picker https://tristen.ca/hcl-picker/#/hlc/6/1/1C1E2A/EFEE69

pal <- c("#1C1E2A","#224653","#21736F","#46A178","#8FCB70","#EFEE69")

pop2000_plot + pop2020_plot +
  plot_layout(guides = "collect") +
  plot_annotation(
    title = "Population density in Africa",
    subtitle = ""
  ) &
  # scale_fill_viridis_c(option = "inferno") &
  scale_fill_gradientn(colors = pal, breaks = seq(0, threshold + 100, 100), labels = c(seq(0, threshold, 100), paste0(">", threshold))) &
  guides(fill = guide_colorbar(barwidth = 30, title.position = "top")) &
  labs(fill = "Population density (humans / km2)") &
  theme_void() &
  theme(
    plot.background = element_rect(fill = "gray20", color = NA),
    legend.position = "bottom",
    text = element_text(family = , color = "gray90" ),
    plot.title = element_text(size = 20, family = "Bauhaus 93"),
    plot.subtitle = element_text(size = 20)
  )
```





